---
title: "Case Study 1- DS 6306"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)
library(tools)
library(sf)
library(naniar)
library(mice)
library (class)

```

# Introduction

| We have undertaken a comprehensive analysis of two important datasets that provide a detailed snapshot of the brewing industry. The first dataset encapsulates a wealth of information on 2,410 U.S. craft beers, covering aspects such as their alcohol by volume (ABV), International Bitterness Units (IBUs), and beer style. This data offers valuable insights into the popularity of various beer styles, the correlation between ABV and IBUs, and potential market trends. The second dataset provides records on 558 U.S. breweries, including their location and size. This dataset will allow us to understand geographical distribution, market saturation, and opportunities for expansion. By cross-referencing these datasets, we can identify patterns and trends that could inform Budweiser's strategic decisions. Our goal is to present these findings in a clear and impactful manner, focusing on the implications of the data rather than the technicalities of the analysis. We trust that these insights will contribute significantly to shaping Budweiser's future strategies. Thank you for your time, and we look forward to discussing our findings with you.

```{r beers and breweries, message=FALSE}
#Import and name each data set
beers <- read.csv("https://github.com/torih1541/Case-Study-1-DDS/files/14513428/Beers.csv")
breweries <- read.csv("https://github.com/torih1541/Case-Study-1-DDS/files/14513429/Breweries.csv")
```

#### How many breweries are present in each state?

| The number of breweries present in each state, sorted by the states with the most breweries: Colorado (CO) - 47 breweries, California (CA) - 39 breweries, Michigan (MI) - 32 breweries, Oregon (OR) - 29 breweries, Texas (TX) - 28 breweries. The list continues with states having fewer breweries, all the way down to those with just one brewery, such as North Dakota (ND), South Dakota (SD), Washington D.C. (DC), and West Virginia (WV).
| This list and map shows a strong presence of breweries in certain states, reflecting perhaps a robust craft beer culture in those locations.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
#Map state abbreviations to their full names
state_map <- c(
  "AL" = "Alabama", AK = "Alaska", AZ = "Arizona", AR = "Arkansas", CA = "California",
  CO = "Colorado", CT = "Connecticut", DE = "Delaware", FL = "Florida", GA = "Georgia",
  HI = "Hawaii", ID = "Idaho", IL = "Illinois", IN = "Indiana", IA = "Iowa",
  KS = "Kansas", KY = "Kentucky", LA = "Louisiana", ME = "Maine", MD = "Maryland",
  MA = "Massachusetts", MI = "Michigan", "MN" = "Minnesota", MS = "Mississippi", MO = "Missouri",
  MT = "Montana", NE = "Nebraska", NV = "Nevada", NH = "New Hampshire", NJ = "New Jersey",
  NM = "New Mexico", NY = "New York", NC = "North Carolina", ND = "North Dakota", OH = "Ohio",
  OK = "Oklahoma", OR = "Oregon", PA = "Pennsylvania", RI = "Rhode Island", SC = "South Carolina",
  SD = "South Dakota", TN = "Tennessee", TX = "Texas", UT = "Utah", VT = "Vermont",
  VA = "Virginia", WA = "Washington", WV = "West Virginia", WI = "Wisconsin", WY = "Wyoming"
)

#Prepare breweries dataset 
breweries <- breweries %>%
  mutate(State = str_trim(State),
         State = toupper(State), 
         StateFullName = state_map[State])

#Aggregate and count number of breweries by state
breweries_by_state <- breweries %>%
  group_by(StateFullName) %>%
  summarise(NumberOfBreweries = n())

#Load US states shapefile for mapping
us_states <- st_read("tl_2023_us_state/tl_2023_us_state.shp")

#Merge the shapefile data with count data
map_data1 <- merge(us_states, breweries_by_state, by.x= "NAME", by.y= "StateFullName", all.x = TRUE)

#Create a map of the nummber of breweries by state
ggplot(data = map_data1) +
  geom_sf(aes(fill = NumberOfBreweries), color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "Number of Breweries") +
  labs(title = "Number of Breweries by State", x = "", y = "") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.15,size = 15),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_rect(color="white"),
        legend.position = 'none',
        plot.margin = unit(c(0, -25, -5, 0), "cm"),
        plot.background = element_rect(color = "white")
        ) + 
        coord_sf(expand= FALSE, datum = NA)

```

#### Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

```{r echo=TRUE}
# Merge the data frames on the Brewery_id (in beers) and Brew_id (in breweries)
merged_data <- merge(beers, breweries, by.x = "Brewery_id", by.y = "Brew_ID")

# Print the first 6 observations
print(head(merged_data, 6))

# Print the last 6 observations
print(tail(merged_data, 6))
```

#### Address the missing values in each column.

The missing ABV values could be missing if the brewery did not or chose not to measure ABV values for some beers and there does not seem to be certain types of beer more likely to be missing ABV so these are missing at random. The large number of missing IBU values could suggest measuring IBU is too expensive or time consuming so breweries would choose not to measure this leaving the values missing at random. We used MCAR test to create a heat map of the data missingness to determine patterns.

```{r}
# Little's MCAR test
mcar_test <- mice::md.pattern(beers)

summary(merged_data)

#Create dataframe with no NA values to perform further analysis
cleaned_data <- na.omit(merged_data)

# Print the first 6 observations
print(head(cleaned_data, 6))

# Print the last 6 observations
print(tail(cleaned_data, 6))
```

#### Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

\*\*\* commentary

```{r}
#Groups cleaned data by state and calculates median ABV and IBU
medians_by_state <- cleaned_data %>%
  group_by(State) %>%
  summarise(Median_ABV = median(ABV, na.rm = TRUE),
            Median_IBU = median(IBU, na.rm = TRUE))

#Bar chart displaying the median ABV for each state
ggplot(medians_by_state, aes(x = reorder(State, Median_ABV), y = Median_ABV)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_hline(yintercept = median(medians_by_state$Median_ABV, na.rm = TRUE), 
             linetype = "dashed", color = "red") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "State", y = "Median ABV (%)", title = "Median Alcohol Content by State") +
  theme(plot.title = element_text(hjust = 0.5))

#Bar chart displaying the median IBU for each state
ggplot(medians_by_state, aes(x = reorder(State, Median_IBU), y = Median_IBU)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  geom_hline(yintercept = median(medians_by_state$Median_IBU, na.rm = TRUE), 
             linetype = "dashed", color = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "State", y = "Median IBU", title = "Median IBU by State") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

```{r paged.print=TRUE}
# Determine which state has the maximum ABV
 state_with_max_ABV <- cleaned_data[which.max(cleaned_data$ABV),]$State
 print(paste("The state with the maximum alcoholic (ABV) beer is", state_with_max_ABV))
 # Determine which state has the maximum IBU
state_with_max_IBU <- cleaned_data[which.max(cleaned_data$IBU),]$State
print(paste("The state with the most bitter (IBU) beer is", state_with_max_IBU))

#Calculate max ABV by state
max_abv_by_state <- cleaned_data %>%
  group_by(State) %>%
  summarise(Max_ABV = max(ABV, na.rm = TRUE))

#Calculate max IBU by state
max_ibu_by_state <- cleaned_data %>%
  group_by(State) %>%
  summarise(Max_IBU = max(IBU, na.rm = TRUE))

#Visualizing Maximum ABV by State
ggplot(max_abv_by_state, aes(x = reorder(State, Max_ABV), y = Max_ABV, fill = State)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = round(Max_ABV, 2)), position = position_stack(vjust = 1.01), hjust = 0.5, size = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "State", y = "Maximum ABV", title = "Maximum ABV by State")


#Visualizing Maximum IBU by State
ggplot(max_ibu_by_state, aes(x = reorder(State, Max_IBU), y = Max_IBU, fill = State)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = round(Max_IBU, 2)), position = position_stack(vjust = 1.01), hjust = 0.5, size = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "State", y = "Maximum IBU", title = "Maximum IBU by State")

```

#### The summary statistics and distribution of the ABV variable

Through an analysis of our beer dataset, we've identified key insights into the Alcohol by Volume (ABV) and International Bitterness Units (IBU). The ABV ranges from 0.1% to 12.8%, with the majority of beers sitting around 5.6%. However, the average ABV is slightly higher at 5.98%, indicating the presence of some significantly stronger beers. When it comes to bitterness as measured by IBUs, the median value is 35, but the mean is 42.71, showing that some beers have a considerably high bitterness level. A regional analysis highlights Kentucky as the producer of the beer with the highest ABV, while Oregon leads with the most bitter beer. It's important to note that these findings are based on our current dataset, and may not represent all beers. We also handled missing data points responsibly; specifically, 62 instances where the ABV was not recorded were excluded from the analysis. These insights could be valuable in shaping marketing strategies and understanding regional taste preferences.

#### Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. Make your best judgment of a relationship and EXPLAIN your answer.

commentary

```{r}
#Scatter plot of relationship between ABV and IBU
ggplot(cleaned_data, aes(x = ABV, y = IBU)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "Alcohol By Volume (ABV)", 
       y = "International Bitterness Units (IBU)", 
       title = "Relationship between Beer Bitterness and Alcoholic Content") +
  geom_smooth(method = "lm", se = FALSE, color = "blue")
```

#### Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other.

\*\*\* KNN & supplementary commentary \*\*\*

```{r}
# Prepare the beers dataset
beers <- beers %>%
  filter(grepl("Ale", Style, ignore.case = TRUE)) %>%
  mutate(Category = ifelse(grepl("IPA", Style, ignore.case = TRUE), "IPA", "Other Ale"))

# Remove missing values
beers <- na.omit(beers)

# Normalize the data
beers$ABV <- scale(beers$ABV)
beers$IBU <- scale(beers$IBU)

# Prepare the data for the KNN classifier
set.seed(5) 
training_index <- sample(1:nrow(beers), 0.8*nrow(beers)) # 80% for training
train <- beers[training_index,]
test <- beers[-training_index,]

# Train the KNN model
knn_model <- knn(train = train[,c('ABV', 'IBU')], 
                 test = test[,c('ABV', 'IBU')],
                 cl = train$Category, 
                 k = 5)

# Validate the model
knn_table <- table(Predicted = knn_model, Actual = test$Category)
knn_table
# Calculate the accuracy
accuracy <- sum(knn_model == test$Category) / length(knn_model)
accuracy 

```

```{r}
# Categorize beers into 'IPA' and 'Other Ale'
beers$Category <- ifelse(grepl("IPA", beers$Style, ignore.case = TRUE), "IPA", 
                         ifelse(grepl("Ale", beers$Style, ignore.case = TRUE), "Other Ale", NA))

# Filter out beers that do not fall into these two categories
beers <- subset(beers, !is.na(Category))

# T-test for ABV differences between IPAs and Other Ales
t_test_abv <- t.test(ABV ~ Category, data = beers, subset = Category %in% c("IPA", "Other Ale"))
t_test_abv

# T-test for IBU differences between IPAs and Other Ales
t_test_ibu <- t.test(IBU ~ Category, data = beers, subset = Category %in% c("IPA", "Other Ale"))
t_test_ibu

# ANOVA for the impact of beer style on ABV
anova_abv <- aov(ABV ~ Category, data = beers)
summary(anova_abv)

# ANOVA for the impact of beer style on IBU
anova_ibu <- aov(IBU ~ Category, data = beers)
summary(anova_ibu)

```

```{r}
# Scatter plot for ABV vs IBU with color-coding for beer types
ggplot(beers, aes(x = IBU, y = ABV, color = Category)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "International Bitterness Units (IBU)", y = "Alcohol By Volume (ABV)", title = "ABV vs IBU by Beer Type") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

# Box plots for ABV and IBU by beer type
ggplot(beers, aes(x = Category, y = ABV, fill = Category)) + 
  geom_boxplot() + 
  labs(x = "Beer Type", y = "Alcohol By Volume (ABV)", title = "ABV Distribution by Beer Type") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")

ggplot(beers, aes(x = Category, y = IBU, fill = Category)) + 
  geom_boxplot() + 
  labs(x = "Beer Type", y = "International Bitterness Units (IBU)", title = "IBU Distribution by Beer Type") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")
```

#### Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence.

```{r}

```

\*\*\* commentary
