---
title: "Case Study 1- DS 6306"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)

```

## Introduction 

The introduction needs to be written as if you are presenting the work to the CEO and CFO of Budweiser (your client) and that they only have had one class in statistics. You may assume that the CEO and CFO gave you the data and gave you the directive to report any interesting finding that you may uncover through your analysis.

```{r beers and breweries}
beers <- read.csv("~/Downloads/Beers.csv")
breweries <- read.csv("~/Downloads/Breweries.csv")
```


How many breweries are present in each state?

```{r}
breweries_by_state <- breweries %>%
  group_by(State) %>%
  summarise(NumberOfBreweries = n())
```

Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

```{r}
# Merge the data frames on the Brewery_id (in beers) and Brew_id (in breweries)
merged_data <- merge(beers, breweries, by.x = "Brewery_id", by.y = "Brew_ID")

# Print the first 6 observations
print(head(merged_data, 6))

# Print the last 6 observations
print(tail(merged_data, 6))
```

Address the missing values in each column.

```{r}
summary(merged_data)
cleaned_data <- na.omit(merged_data)

```

Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

```{r}
medians_by_state <- cleaned_data %>%
  group_by(State) %>%
  summarise(Median_ABV = median(ABV, na.rm = TRUE),
            Median_IBU = median(IBU, na.rm = TRUE))

ggplot(medians_by_state, aes(x = State)) +
  geom_bar(aes(y = Median_ABV, fill = "Median ABV"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = Median_IBU, fill = "Median IBU"), stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Median Value", fill = "Metric") +
  ggtitle("Median Alcohol Content and IBU by State")
```

Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

```{r}

```

Comment on the summary statistics and distribution of the ABV variable.

Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

```{r}

```

